Computing astrophysical inferences with diffusive nested sampling

### 2014-01-28 ms.tex
Intro work

\section{Introduction}
The major goal of astronomy, as with most sciences, is to understand the
universe as well as we can while recognising that we will never have all of
the relevant information that we might want.

In many subfields within astronomy, Bayesian inference has (rightfully)
become the dominant framework for analysing data and quantifying the
plausibility of hypotheses about the universe \citep[e.g.][]{refs}.

The first step is to construct a set of possible answers to the question of
interest, that is, the ``hypothesis space'', indexed by parameters $\pars$.
A space of possible data sets $\data$ is considered, and a joint prior is
assigned:
\begin{eqnarray}
p(\pars, \data) = p(\pars) p(\data | \pars) = p(\data)p(\pars | \data)
\end{eqnarray}
The marginal distribution $p(\pars)$, which is usually called the ``prior'',
describes initial uncertainty about the value of the parameters. When the
data $\data$ becomes known, our state of knowledge about $\pars$ changes
from the marginal $p(\pars)$ to the conditional distribution $p(\pars | \data)$
that corresponds to the observed data, usually called the ``posterior''.
Incidentally, Bayesian inference can be understood as a special case of
MaxEnt updating, starting from the joint prior $p(\pars, \data)$ and applying
the constraint $P(\data = \data_{\rm actual}) = 1$ \citep{caticha}. Note that
this is not a ``derivation'' of Bayesian updating as the rules of probability
already imply a commitment to update to $p(\pars | \data)$ once $\data$ becomes
known (what else could a conditional probability be?).

>>>>

\section{Introduction}
The major goal of astronomy, as with most sciences, is to understand the
universe as well as we can while recognising that we will never have all of
the relevant information that we might want. Probability theory is the
appropriate framework for describing and quantifying uncertainty
\citep{cox, jaynes}. In recent years, Bayesian inference has (rightfully)
become the dominant framework for analysing data and quantifying the
plausibility of hypotheses about the universe, and has displaced
more ad-hoc methodologies in many subfields of
astronomy \citep[e.g.][]{examples}.

Bayesian inference is usually performed by
constructing a set of possible answers to the question of
interest, that is, the ``hypothesis space'', indexed by parameters $\pars$.
A state of prior knowledge about the parameters is described by assigning a
prior distribution $p(\pars)$. Some relevant data $\data$ is about to become
known, and prior knowledge about a connection between $\theta$ and $\data$ is
expressed by a choice of ``sampling distribution'' $p(\data | \theta)$, which
is really a family of probability distributions, one for each possible value of
$\theta$.

By the product rule, the joint (prior) distribution is
\begin{eqnarray}
p(\pars, \data) = p(\pars) p(\data | \pars) = p(\data)p(\pars | \data)
\end{eqnarray}
When the
data $\data$ becomes known, our state of knowledge about $\pars$ changes
from the marginal $p(\pars)$ to the conditional distribution
$p(\pars | \data_{\rm actual})$
that corresponds to the observed data. The distribution
$p(\pars | \data_{\rm actual})$
called the ``posterior''.
Incidentally, Bayesian inference can be understood as a special case of
MaxEnt updating, starting from the joint prior $p(\pars, \data)$ and applying
the constraint $P(\data = \data_{\rm actual}) = 1$ \citep{caticha}. Note that
this is not a ``derivation'' of Bayesian updating as the rules of probability
already imply a commitment to update from $p(\theta)$ to $p(\theta | \data)$
once $\data$ becomes known.

Computing the consequences of a posterior distribution can be a challenging
task. Usually we want to marginalise out nuisance parameters, to compute
particular posterior probabilities such as $P(\theta > 1.5)$, or to summarise
the posterior distribution for easy communication with others.
Markov Chain Monte Carlo (MCMC) algorithms provide a general set of tools
for achieving these goals.
